---
title: "models"
output: html_document
---
#read in

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in files}


all_data=read.csv("C:/Users/torda/Documents/egyetem/emoji/temporary data/all_data_pca.csv")

study1=read.csv("C:/Users/torda/Documents/egyetem/emoji/temporary data/study1_data_pca.csv")
study2=read.csv("C:/Users/torda/Documents/egyetem/emoji/temporary data/study2_data_pca.csv")

study1$emoji= factor(study1$emoji, levels= c("no_emoji", "face_emoji"))

study1$education= factor(study1$education, levels= c("less than primary school", "primary school", "secondary school", "bachelor", "master", "PhD", "other"))

study2$education= factor(study2$education, levels= c("less than primary school", "primary school", "secondary school", "bachelor", "master", "PhD", "other"))

```

#hypotheses
Study 1. Face-emoji

Hypothesis 1. Using face emoji in an informal online environment helps to cause a better first impression of the sender of the message by raising perceived warmth of the person while not affecting their perceived competence.

Hypothesis 2. Using face emoji in a formal online environment causes a worse first impression of the sender of the message as it does not help to raise perceived warmth of the person while it decreases their perceived competence.

Hypothesis 3. In a Danish formal email with emoji the perceived competence of the sender will be higher than the one in a Hungarian formal email with emoji.


Study 2. Non-face emoji #same as for study1
Hypothesis 1. Using non-face emoji in an informal online environment helps to cause a better first impression of the sender of the message by raising perceived warmth of the person while not affecting their perceived competence.

Hypothesis 2. Using non-face emoji in a formal online environment causes a worse first impression of the sender of the message as it does not help to raise perceived warmth of the person while it decreases their perceived competence.

Hypothesis 3. The perceived competence of the sender will be higher in a Danish formal email with emoji than the one in a Hungarian formal email with emoji.

Exploratory analysis:
Does perceived gender matter for perceived warmth and competence in any of the conditions?

Maybe also explore the difference between face and non-face emoji, given that I can?

#study 1

Does the presence of the emoji matter for first impression, given the formality of the text?
-	warmth ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality
- competence ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality
```{r }
library(lmerTest)
#hypothesis 1/a: higher warmth for informal
model1= lmer(warmness ~ emoji + formality + emoji*formality + (1|id), study1)
#hypothesis 1/b: same competence for informal and formal
model2= lmer(competence ~ emoji + formality + emoji*formality + (1|id), study1)

#hypothesis 2/a: same wamrth for formal and informal
#?
#hypothesis 2/b: lower competence for formal

```

For formal data with emoji look at difference in culture
-	expecting Danish to have a higher competence than Hungarian
-	competence ~ α + β1 *emoji + β2 *country + β3 *emoji*formality*culture 

```{r }

#hypothesis 3: higher competence for danish
model5= lmer(competence ~ emoji + culture + emoji*formality*culture + (1|id), study1)


```

Appropriateness
-	Does the rating of appropriateness explain the ratings for warmth and competence?
-	I am not sure how to do this.
- I'm adding appro. to the interection model because I want to see if it explains variance in the data that is not explained by the interaction.
- Maybe it should be an interaction, right? I don't know :(
-	warmth ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality + β4* appropriateness
-	competence ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality + β4* appropriateness

```{r }

model6= lmer(warmness ~ emoji + formality + emoji*formality + appropriateness + (1|id), study1)

model7= lmer(competence ~ emoji + formality + emoji*formality + appropriateness + (1|id), study1)
```

Exploratory analysis – gender of the sender
Does perceived gender matter for perceived warmth and competence in any of the conditions?
-	warmth ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality + β4*sender_gender
-	competence ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality + β4*sender_gender

```{r }

model8= lmer(warmness ~ emoji + formality + emoji*formality + sender_gender + (1|id), study1)

model9= lmer(competence ~ emoji + formality + emoji*formality + sender_gender + (1|id), study1)
```

Exploratory analysis – emoji usage
Does the self-reported frequency of using emoji matter for perceived warmth and competence in any of the conditions?
-	warmth ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality + β4*emoji_usage
-	competence ~ α + β1 *emoji + β2 *formality + β3 *emoji*formality + β4*emoji_usage

```{r }

model10= lmer(warmness ~ emoji + formality + emoji*formality + emoji_usage + (1|id), study1)

model11= lmer(competence ~ emoji + formality + emoji*formality + emoji_usage + (1|id), study1)
```

#study 2

Same...

